---
title: "Homework_2 - Bayesian Statistics"
author: "Domagoj Korais"
date: "Spring 2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(message=FALSE)
library(tidyverse)
library(arm)
set.seed(pi)
```

#Ex01
```{r ex01}
library(expm)
#build the transition matrix
dim=6
diagonalValue=0.5
offDiagonalValue=0.25
P=diag(diagonalValue, dim, dim)


for (row in 1:dim) {
  for (col in 1:dim) {
    if (row==col + 1 || col == row + 1 ) {
      P[row,col]=0.25
    }  
  }
}
#special cases
P[1,2] = diagonalValue
P[6,5] = diagonalValue
#starting vector
val=1/dim
start=rep(val,dim)
#compute stationary state:

final=start %*% (P %^% 30)
print(final)  


```
As we can see the final probability distibution converges to the desired one.

# Ex03
```{r ex03 }
#Random walk


randomWalk = function(N,d){
  x=rep(0,N)
  count = 0
  for (i in 2:N) {
    y = x[i-1]+runif(1,-d,d)
    prob = min(1,exp( (x[i-1]^2-y^2) / 2 )) 
    u=runif(1)
    
    if(prob>u){
      count = count +1
      
      x[i]=y  
    }
    else{
      x[i]=x[i-1]
    }
  }
  accRatio=count/N
  return(list(x,accRatio))
}
d=5000
low=d-500
a=randomWalk(d,0.1)
mainA="Uniform 0.1"
b=randomWalk(d,1)
mainB="Uniform 1"
c=randomWalk(d,10)
mainC="Uniform 10"
par(mfrow=c(1,3))
print("Uniform case acceptance ratio: ")
print(c(a[[2]],b[[2]],c[[2]]))
plot(a[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainA)
plot(b[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainB)
plot(c[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainC)
hist(a[[1]], breaks=100, border="gray40",freq=F, main=mainA,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
hist(b[[1]], breaks=100, border="gray40",freq=F, main = mainB,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
hist(c[[1]], breaks=100, border="gray40",freq=F, main= mainC,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
acf(a[[1]], main=mainA)
acf(b[[1]], main=mainB)
acf(c[[1]], main=mainC)
library(distr)
randomWalkLaplace = function(N,lambda){
  x=rep(0,N)
  count=0
  for (i in 2:N) {
    y = x[i-1]+r(DExp(lambda))(1)
    prob = min(1,exp( (x[i-1]^2-y^2) / 2 )) 
    u=runif(1)
    
    if(prob>u){
      count = count +1
      x[i]=y  
    }
    else{
      x[i]=x[i-1]
    }
  }
   accRatio=count/N
  return(list(x,accRatio))
}

d=5000
low=d-500
a=randomWalkLaplace(d,0.1)
mainA="Laplace 0.1"
b=randomWalkLaplace(d,1)
mainB="Laplace 1"
c=randomWalkLaplace(d,10)
mainC="Laplace 10"
par(mfrow=c(1,3))
print("Laplace case acceptance ratio: ")
print(c(a[[2]],b[[2]],c[[2]]))
plot(a[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainA)
plot(b[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainB)
plot(c[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainC)
hist(a[[1]], breaks=100, border="gray40",freq=F, main=mainA,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
hist(b[[1]], breaks=100, border="gray40",freq=F, main = mainB,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
hist(c[[1]], breaks=100, border="gray40",freq=F, main= mainC,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
acf(a[[1]], main=mainA)
acf(b[[1]], main=mainB)
acf(c[[1]], main=mainC)

randomWalkCauchy = function(N,scale){
  x=rep(0,N)
  count=0
  for (i in 2:N) {
    y = x[i-1]+rcauchy(1,0,scale)
    prob = min(1,exp( (x[i-1]^2-y^2) / 2 )) 
    u=runif(1)
   
    if(prob>u){
      count = count +1
      x[i]=y  
    }
    else{
      x[i]=x[i-1]
    }
  }
   accRatio=count/N
  return(list(x,accRatio))
}

d=5000
low=d-500
a=randomWalkCauchy(d,0.1)
mainA="Cauchy 0.1"
b=randomWalkCauchy(d,1)
mainB="Cauchy 1"
c=randomWalkCauchy(d,10)
mainC="Cauchy 10"
par(mfrow=c(1,3))
print("Cauchy case acceptance ratio: ")
print(c(a[[2]],b[[2]],c[[2]]))
plot(a[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainA)
plot(b[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainB)
plot(c[[1]][low:d], type="l", ylab="Trace", xlab="Iteration",main = mainC)
hist(a[[1]], breaks=100, border="gray40",freq=F, main=mainA,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
hist(b[[1]], breaks=100, border="gray40",freq=F, main = mainB,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
hist(c[[1]], breaks=100, border="gray40",freq=F, main= mainC,xlim = c(-4,4))
curve(dnorm(x),add=TRUE,col="red")
acf(a[[1]], main=mainA)
acf(b[[1]], main=mainB)
acf(c[[1]], main=mainC)

```

Comments:

a) From the autocorrelation point of view the  best choice looks to be the Uniform 01 since the autocorrelation function decreases faster.

b) The choice of the distribution looks quite unimportant to me, much more important is an optimal choice of parameters.

c) Let's do some grid search of the hyperparameters:

As we can check from the computed values To reach a 25% acceptance ratio the values for the distributions are the following: For the uniform a value between 5 and 6, For Cauchy a value between 3 and 4 and for Laplace a value between 0.2 and 0.3.
```{r}
#get a 0.25 acc ratio
N=1000
for (i in 1:10) {
  print(paste0("value: ", i))
  print(randomWalk(N,i)[[2]])
  print(randomWalkCauchy(N,i)[[2]])
  print(randomWalkLaplace(N,i)[[2]])
  
}
for (i in 1:10) {
  j=i*0.1
  print(paste0("value: ", j))
  print(randomWalk(N,j)[[2]])
  print(randomWalkCauchy(N,j)[[2]])
  print(randomWalkLaplace(N,j)[[2]])
  
}

```

