---
title: "exam"
author: "Domagoj Korais"
date: "June 17, 2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(message=FALSE)
library(tidyverse)

set.seed(pi)
```

# Dataset description

This dataset describes the mortality rate of malignent melanoma in eight European countries.
The reference for the used dataset is the following: 

<blockquote>
<p>Langford, I.H., Bentham, G. and McDonald, A. (1998). Multilevel modelling of geographically aggregated health data: a case study on malignant melanoma mortality and UV exposure in the European community. <em>Statistics in Medicine</em>, <em>17</em>, 41-58.</p>
</blockquote>

There is a total of six variables:
<ul>
<li><code>nation</code> a factor with levels <code>Belgium</code>, <code>W.Germany</code>, <code>Denmark</code>, <code>France</code>, <code>UK</code>, <code>Italy</code>, <code>Ireland</code>, <code>Luxembourg</code>, and <code>Netherlands</code></li>
<li><code>region</code> Region ID - a factor</li>
<li><code>county</code> County ID - a factor</li>
<li><code>deaths</code> Number of male deaths due to malignant melanoma during 1971-1980</li>
<li><code>expected</code> Number of expected deaths</li>
<li><code>uvb</code> Centered measure of the UVB dose reaching the earth’s surface in each county.</li>
</ul>

The data are geographically nested, since nations are divided in regions, and regions are divided in counties. 
A county is the smallest area forwich we have data.
To be able to compare different nations I decided to include population data of the countries, this made possible to compare death ratio instead of raw counts.


## Exploratory data analysis
```{r eda}
library(mlmRev)
library(tidyverse)
library(GGally)
data = Mmmec

globalNationData = data%>%
  group_by(nation)%>%
  summarize(total=sum(deaths),
            total_expected=sum(expected),
            mean_uvb=mean(uvb),
            sd_uvb = sd(uvb))%>%
  mutate(delta=total-total_expected)%>%
  arrange(desc(delta))
#plot
ggplot(data,aes(x=nation,y=uvb))+
  geom_boxplot()+
  ggtitle("Uvb in nations")

ggplot(data,aes(x=nation,y=deaths))+
  geom_boxplot()+
  ggtitle("deaths in nations")


summarise(data)

#plot global data
ggplot(data = globalNationData,aes(x=mean_uvb,y=delta,color=nation))+
  geom_point()+
  ggtitle("Residuals vs mean uvb at national level ")

globalRegionData = data%>%
  group_by(region)%>%
  summarize(total=sum(deaths),
            total_expected=sum(expected),
            mean_uvb=median(uvb),
            sd_uvb = sd(uvb))%>%
  mutate(delta=total-total_expected)%>%
  arrange(desc(delta))
globalRegionData = inner_join(globalRegionData,data,"region")
#plot global  data
ggplot(data = globalRegionData,aes(x=mean_uvb,y=delta,color = nation))+
  geom_point()+
  ggtitle("Residuals vs median uvb at regional level ")

ggplot(data = globalRegionData,aes(x=mean_uvb,y=total,color = nation))+
  geom_point()+
  ggtitle("Total deaths vs median uvb at regional level ")

topRegions = data%>%group_by(region)%>%summarize(total=sum(deaths),count = n())%>%arrange(desc(total))%>%top_n(10,wt=total)



data%>%group_by(county)%>%count() #no repetitions
totalDeaths = data%>%summarize(total=sum(deaths))
#countries with top regions
inner_join(data,topRegions,by = "region")%>%arrange(desc(total))
#number of deaths from top regions

ggplot(data,aes(x=uvb,y=deaths,color=nation))+
  geom_point()

```

```{r summary of the dataset}
data_complete2%>%group_by(nation)%>%summarise(total_deaths =sum(deaths), deaths_mean=mean(deaths),deaths_sd=sd(deaths),expected_mean=mean(expected),male_population=mean(male_population))
```

# Integrate with population of country
The population data are taken from the following dataset and are related to the 1980 year: 
https://openei.org/doe-opendata/dataset/a7fea769-691d-4536-8ed3-471e993a2445
In order to get population data on the county level I choose to use the values of the expected deaths, this enabled me to rescale the values from the nation-level ones. According to my analysis this operation clearly improves the model performance, since otherwise it would be not possible to compare the counts of different regions. The population variable is trathed in the following as an exposure variable.

```{r population}

populationbycountry19802010millions <- read_csv("populationbycountry19802010millions.csv", 
    col_types = cols(`1980` = col_double()))

#rename Western Germany to make in compatible with our data
filter(populationbycountry19802010millions,X1 =="Germany, West")

populationbycountry19802010millions = populationbycountry19802010millions%>%mutate(X1 =replace(X1,X1=="Germany, West","W.Germany"))

pop = populationbycountry19802010millions%>%select(nation=X1,population = '1980')

data_complete = inner_join(pop,data,by=c("nation"="nation"))
summary(data_complete)
sd(data_complete$deaths)
```

#Let's compute the death ratio
```{r death_ratio}
data_complete=mutate(data_complete,ratio = (deaths)/population,residuals = (deaths-expected)/population)
#And now let's rock with ratios
ggplot(data_complete,aes(x=nation,y=ratio,color=nation))+
  geom_boxplot()

ggplot(data_complete,aes(x=uvb,y=log2(ratio),color=nation))+
  geom_point()


dataReduced = data_complete %>%select(-c(region,county))
ggpairs( dataReduced ) 


```
# Compute the death ratio for region
```{r }
nation = data%>%group_by(nation)%>%summarise(tot_expected=sum(expected))
nation_complete = inner_join(pop,nation,by=c("nation"="nation"))
male_nation_complete = nation_complete%>%mutate(male_population=population/2)%>%select(-population)
data_complete2 = inner_join(male_nation_complete,data,by=c("nation"="nation"))
data_complete2 = data_complete2%>%mutate(local_male_population=(male_population*expected)/tot_expected)
data_complete2 = data_complete2%>%mutate(local_rate = (deaths)/(local_male_population))
ggplot(data_complete2,aes(x=local_rate))+
  geom_histogram()

ggplot(data_complete2,aes(x=deaths))+
  geom_histogram()
dataReduced = data_complete2 %>%select(-c(region,county))
ggpairs( dataReduced ) 
```

```{R}
#Let's try a Poisson regression with ungrouped data, using nation as a regressor for rate
library(rstanarm)
library(bayesplot) 

fit <- stan_glm(deaths ~ uvb + nation + region , data=data_complete, family=neg_binomial_2(),cores=2)
saveRDS(fit,"models/fit.stanModel")
a=readRDS("models/fit.stanModel")
summary(a)
posterior <- as.array(fit)
dim(posterior)
mcmc_areas(posterior)
mcmc_intervals(posterior)
mcmc_hist(posterior)
mcmc_scatter(posterior, pars = c("(Intercept)","nationItaly"))
mcmc_trace(posterior)
mcmc_parcoord(posterior)
mcmc_pairs(posterior)
yrep_poisson <- posterior_predict(fit, draws = 500)
y=data_complete$deaths
y
data_complete$deaths
ppc_stat(y,yrep_poisson)
ppc_stat_grouped(y,yrep_poisson,group=data_complete$nation)+
  ggtitle("Using raw counts ")

ppc_stat_grouped(y,yrep_poisson,group=data_complete$nation, stat="sd")+
  ggtitle("Using raw counts sd comparison")
ppc_dens_overlay(y,yrep_poisson)
dim(y)
```


```{R}
#Let's try a Poisson regression with ungrouped data, using nation as a regressor for rate
library(rstanarm)
library(bayesplot) 

fit3 <- stan_glm(deaths ~ uvb + nation, data=data_complete, family=neg_binomial_2(),cores=2)
summary(fit3)
posterior <- as.array(fit3)
dim(posterior)
mcmc_areas(posterior)
mcmc_intervals(posterior)
mcmc_hist(posterior)
mcmc_scatter(posterior, pars = c("(Intercept)","nationItaly"))
mcmc_trace(posterior)
mcmc_parcoord(posterior)
mcmc_pairs(posterior)
yrep_poisson <- posterior_predict(fit3, draws = 500)
y=data_complete$deaths
y
data_complete$deaths
ppc_stat(y,yrep_poisson)
ppc_stat_grouped(y,yrep_poisson,group=data_complete$nation)+
  ggtitle("Using raw counts ")

ppc_stat_grouped(y,yrep_poisson,group=data_complete$nation, stat="sd")+
  ggtitle("Using raw counts sd comparison")
ppc_dens_overlay(y,yrep_poisson)
dim(y)

#residuals





```
```{R}
#Let's try a Poisson regression with ungrouped data, using nation as a regressor for rate on the rates!
library(rstanarm)
library(bayesplot) 

fit2 <- stan_glm(deaths ~ uvb + nation + offset(male_population) , data=data_complete2, family=neg_binomial_2(),cores=2)
saveRDS(fit2,"models/fit2.stanModel")
summary(fit2)
posterior <- as.array(fit2)
dim(posterior)
mcmc_areas(posterior)
mcmc_intervals(posterior)
mcmc_hist(posterior)
mcmc_scatter(posterior, pars = c("(Intercept)","nationItaly"))
mcmc_trace(posterior)
mcmc_parcoord(posterior)
mcmc_pairs(posterior)
yrep_poisson <- posterior_predict(fit2, draws = 500)
y=data_complete2$deaths
y
data_complete$deaths
ppc_stat(y,yrep_poisson)
ppc_stat_grouped(y,yrep_poisson,group=data_complete$nation)+
  ggtitle("Using exposures means comparison")

ppc_stat_grouped(y,yrep_poisson,group=data_complete$nation, stat="sd")+
  ggtitle("Using exposures sd comparison")
ppc_dens_overlay(y,yrep_poisson)
dim(y)
```

```{r}

compare_fits = function (fit,title_mean,title_sd,title_overlay,data){
  posterior <- as.array(fit)
  yrep_poisson <- posterior_predict(fit, draws = 500)
  y=data$deaths
  
  # ppc_stat_grouped(y,yrep_poisson,group=data$nation)+
  #   ggtitle(title_mean)

  #ppc_stat_grouped(y,yrep_poisson,group=data$nation, stat="sd")+
  #  ggtitle(title_sd)
  # 
  # ppc_dens_overlay(y,yrep_poisson)+
  #   ggtitle(title_overlay)
  #  ppc_stat_2d(y,yrep_poisson)+
   # ggtitle(title_sd)
    
    mean_y_rep <- colMeans(yrep_poisson)
std_resid <- (y - mean_y_rep) / sqrt(mean_y_rep)
qplot(mean_y_rep, std_resid) + hline_at(2) + hline_at(-2)+
  labs(x="Mean of y_rep", y= "Stand. residuals")+
  xaxis_text(on =TRUE, size=22)+
  yaxis_text(on =TRUE, size=22)+
    theme(axis.title.x = element_text(size=16),
          axis.title.y = element_text(size=16))
  
}

compare_fits(fit,"raw count mean","raw count sd","raw overlay",data=data_complete2)
compare_fits(fit2,"exposure count mean","esposure count sd","exposure overlay",data=data_complete2)
compare_fits(fit3,"only uvb mean","only uvb count sd","only uvb overlay",data=data_complete2)
  
```
