---
title: "exam"
author: "Domagoj Korais"
date: "June 17, 2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(message=FALSE)
library(tidyverse)

set.seed(pi)
```

# Dataset description

This dataset describes the mortality rate of malignent melanoma in a bunch of European countries. The reference for the used dataset is the following: 

<blockquote>
<p>Langford, I.H., Bentham, G. and McDonald, A. (1998). Multilevel modelling of geographically aggregated health data: a case study on malignant melanoma mortality and UV exposure in the European community. <em>Statistics in Medicine</em>, <em>17</em>, 41-58.</p>
</blockquote>

There is a total of six variables:
<ul>
<li><code>nation</code> a factor with levels <code>Belgium</code>, <code>W.Germany</code>, <code>Denmark</code>, <code>France</code>, <code>UK</code>, <code>Italy</code>, <code>Ireland</code>, <code>Luxembourg</code>, and <code>Netherlands</code></li>
<li><code>region</code> Region ID - a factor</li>
<li><code>county</code> County ID - a factor</li>
<li><code>deaths</code> Number of male deaths due to malignant melanoma during 1971-1980</li>
<li><code>expected</code> Number of expected deaths</li>
<li><code>uvb</code> Centered measure of the UVB dose reaching the earth’s surface in each county.</li>
</ul>

The data are geographically hierarchical, since nations are divided in regions, and regions are divedid in counties. 


## Exploratory data analysis
```{r eda}
library(mlmRev)
library(tidyverse)
data = Mmmec

globalNationData = data%>%
  group_by(nation)%>%
  summarize(total=sum(deaths),
            total_expected=sum(expected),
            mean_uvb=mean(uvb),
            sd_uvb = sd(uvb))%>%
  mutate(delta=total-total_expected)%>%
  arrange(desc(delta))
#plot
ggplot(data,aes(x=nation,y=uvb))+
  geom_boxplot()+
  ggtitle("Uvb in nations")




#plot global data
ggplot(data = globalNationData,aes(x=mean_uvb,y=delta,color=nation))+
  geom_point()+
  ggtitle("Residuals vs mean uvb at national level ")

globalRegionData = data%>%
  group_by(region)%>%
  summarize(total=sum(deaths),
            total_expected=sum(expected),
            mean_uvb=median(uvb),
            sd_uvb = sd(uvb))%>%
  mutate(delta=total-total_expected)%>%
  arrange(desc(delta))
globalRegionData = inner_join(globalRegionData,data,"region")
#plot global  data
ggplot(data = globalRegionData,aes(x=mean_uvb,y=delta,color = nation))+
  geom_point()+
  ggtitle("Residuals vs median uvb at regional level ")

ggplot(data = globalRegionData,aes(x=mean_uvb,y=total,color = nation))+
  geom_point()+
  ggtitle("Total deaths vs median uvb at regional level ")

topRegions = data%>%group_by(region)%>%summarize(total=sum(deaths),count = n())%>%arrange(desc(total))%>%top_n(10,wt=total)



data%>%group_by(county)%>%count() #no repetitions
totalDeaths = data%>%summarize(total=sum(deaths))
#countries with top regions
inner_join(data,topRegions,by = "region")%>%arrange(desc(total))
#number of deaths from top regions

ggplot(data,aes(x=uvb,y=deaths,color=nation))+
  geom_point()



```

```{r population}

populationbycountry19802010millions <- read_csv("/home/doma/Downloads/populationbycountry19802010millions.csv")
#rename Western Germany
filter(populationbycountry19802010millions,X1 =="Germany, West")

populationbycountry19802010millions = populationbycountry19802010millions%>%mutate(X1 =replace(X1,X1=="Germany, West","W.Germany"))
population = populationbycountry19802010millions%>%rename(nation=X1,population = '1980')%>%select(nation,population)
data = inner_join(population,data,by=c("nation"="nation"))

```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
