---
title: "exam"
author: "Domagoj Korais"
date: "June 17, 2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(tidyverse)
library(gridExtra)

set.seed(pi)
```

# Dataset description

This dataset describes the mortality rate of malignent melanoma in eight European countries.
The reference for the used dataset is the following: 

<blockquote>
<p>Langford, I.H., Bentham, G. and McDonald, A. (1998). Multilevel modelling of geographically aggregated health data: a case study on malignant melanoma mortality and UV exposure in the European community. <em>Statistics in Medicine</em>, <em>17</em>, 41-58.</p>
</blockquote>

There is a total of six variables:
<ul>
<li><code>nation</code> a factor with levels <code>Belgium</code>, <code>W.Germany</code>, <code>Denmark</code>, <code>France</code>, <code>UK</code>, <code>Italy</code>, <code>Ireland</code>, <code>Luxembourg</code>, and <code>Netherlands</code></li>
<li><code>region</code> Region ID - a factor</li>
<li><code>county</code> County ID - a factor</li>
<li><code>deaths</code> Number of male deaths due to malignant melanoma during 1971-1980</li>
<li><code>expected</code> Number of expected deaths</li>
<li><code>uvb</code> Centered measure of the UVB dose reaching the earth’s surface in each county.</li>
</ul>

The data are geographically nested, since nations are divided in regions, and regions are divided in counties. 
A county is the smallest area forwich we have data.
To be able to compare different nations I decided to include population data of the countries, this made possible to compare death ratio instead of raw counts.


## Exploratory data analysis
```{r eda}
library(mlmRev)
library(tidyverse)
library(GGally)
library(knitr)
data = Mmmec

globalNationData = data%>%
  group_by(nation)%>%
  summarize(total=sum(deaths),
            total_expected=sum(expected),
            mean_uvb=mean(uvb),
            sd_uvb = sd(uvb))%>%
  mutate(delta=(total-total_expected)/total)%>%
  arrange(desc(delta))
#plot
p1 = ggplot(data,aes(x=nation,y=uvb))+
  geom_boxplot()+
  ggtitle("Uvb in nations")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

p2 = ggplot(data,aes(x=nation,y=deaths))+
  geom_boxplot()+
  ggtitle("deaths in nations")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

grid.arrange(p1, p2, ncol=2)

p3 = ggplot(data,aes(x=deaths))+
  geom_histogram()+
  ggtitle("Deaths distribution")
  

p4 = ggplot(data,aes(x=uvb))+
  geom_histogram()+
  ggtitle("Uvb distribution")

grid.arrange(p3, p4, ncol=2)

#residuals at county level
data%>%
  mutate(delta=(deaths-expected)/deaths)%>%
  ggplot(aes(x=deaths,y=delta,color = nation))+
  geom_point()+
  ggtitle("Std residuals vs mean uvb at county level  ")




# globalRegionData = data%>%
#   group_by(region)%>%
#   summarize(total=sum(deaths),
#             total_expected=sum(expected),
#             mean_uvb=median(uvb),
#             sd_uvb = sd(uvb))%>%
#   mutate(delta=(total-total_expected)/total)%>%
#   arrange(desc(delta))
# globalRegionData = inner_join(globalRegionData,data,"region")
# #plot global  data
# ggplot(data = globalRegionData,aes(x=total,y=delta,color = nation))+
#   geom_point()+
#   ggtitle("Residuals vs mean uvb at regional level ")
# 
# 
# 
# ggplot(data = globalRegionData,aes(x=mean_uvb,y=total,color = nation))+
#   geom_point()+
#   ggtitle("Total deaths vs median uvb at regional level ")
# 
# topRegions = data%>%group_by(region)%>%summarize(total=sum(deaths),count = n())%>%arrange(desc(total))%>%top_n(10,wt=total)
# 
# 
# 
# data%>%group_by(county)%>%count() #no repetitions
# totalDeaths = data%>%summarize(total=sum(deaths))
# #countries with top regions
# inner_join(data,topRegions,by = "region")%>%arrange(desc(total))
# #number of deaths from top regions
# 
# ggplot(data,aes(x=uvb,y=deaths,color=nation))+
#   geom_point()

```



# Integration with population information
The population data are taken from the following dataset and are related to the 1980 year: 
https://openei.org/doe-opendata/dataset/a7fea769-691d-4536-8ed3-471e993a2445
In order to get population data on the county level I choose to use the values of the expected deaths, this enabled me to rescale the values from the nation-level ones. According to my analysis this operation clearly improves the model performance, since otherwise it would be not possible to compare the counts of different regions. The population variable is trathed in the following as an exposure variable.

```{r population}

populationbycountry19802010millions <- read_csv("populationbycountry19802010millions.csv", 
    col_types = cols(`1980` = col_double()))

#rename Western Germany to make in compatible with our data
populationbycountry19802010millions = populationbycountry19802010millions%>%mutate(X1 =replace(X1,X1=="Germany, West","W.Germany"))

pop = populationbycountry19802010millions%>%select(nation=X1,population = '1980')

data_complete = inner_join(pop,data,by=c("nation"="nation"))

#sd(data_complete$deaths)
```

```{r summary of the dataset}
nation = data%>%group_by(nation)%>%summarise(tot_expected=sum(expected))
nation_complete = inner_join(pop,nation,by=c("nation"="nation"))
male_nation_complete = nation_complete%>%mutate(male_population=population/2)%>%select(-population)
data_complete2 = inner_join(male_nation_complete,data,by=c("nation"="nation"))
data_complete2 = data_complete2%>%mutate(local_male_population=(male_population*expected)/tot_expected)
data_complete2 = data_complete2%>%mutate(local_rate = (deaths)/(local_male_population))

kable(data_complete2%>%group_by(nation)%>%summarise(total_deaths =sum(deaths), deaths_mean=mean(deaths),deaths_sd=sd(deaths),expected_mean=mean(expected),male_population=mean(male_population)),caption = "Summary at national level")
```


#Let's compute the death ratio
```{r death_ratio}
data_complete=mutate(data_complete,ratio = (deaths)/population,residuals = (deaths-expected)/population)
#And now let's rock with ratios
ggplot(data_complete,aes(x=nation,y=ratio,color=nation))+
  geom_boxplot()

ggplot(data_complete,aes(x=uvb,y=log2(ratio),color=nation))+
  geom_point()


dataReduced = data_complete %>%select(-c(region,county))
ggpairs( dataReduced ) 


```

#Models comparison

```{r compute comparison}
library(loo)
library(bayesplot)
#function for loo from stan model
custom_loo = function(fitted_stan_model){
  log_lik_slopes = extract_log_lik(fitted_stan_model)
  loo_slopes <- loo(log_lik_slopes)
  return(loo_slopes)
}
#import models, the models are the ones created using the file model_factory
fit1=readRDS("models/fit1.stanModel")
fit2=readRDS("models/fit2.stanModel")
fit3=readRDS("models/fit3.stanModel")
fit4=readRDS("models/fit4.stanModel")
fit5=readRDS("models/fit5.stanModel")
fit6=readRDS("models/fit6.stanModel")
fit7=readRDS("models/fit7.stanModel")
fit9=readRDS("models/fit9.stanModel")
fit10=readRDS("models/fit10.stanModel")
fit11=readRDS("models/fit11.stanModel")
fit12=readRDS("models/fit12.stanModel")
#compare models
loo_1  = loo(fit1)
loo_2  = loo(fit2)
loo_3  = loo(fit3)
loo_4  = loo(fit4)
loo_5  = loo(fit5)
loo_6  = loo(fit6)
loo_7  = loo(fit7)
loo_9  = custom_loo(fit9)
loo_10  = custom_loo(fit10)
loo_11  = custom_loo(fit11)
loo_12  = custom_loo(fit12)
loo_diff = compare(loo_1,loo_2,loo_3,loo_4,loo_5,loo_6,loo_7,loo_9,loo_10,loo_11,loo_12)

```

```{r show comparison}
kable(loo_diff,caption="Comparison between models")
```